Secure Composition of Robust and Optimising Compilers

### 2023-07-08 popl/main.tex
Add cits

\begin{itemize}
  \item %
        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\MP{cits}).
        % 
        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose both their formalisation and their composition.
        % 
        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
        % 
        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\MP{cits}.
        \MPin{
        	conclude that did this prop in a single pass, we show how to do it compositionally
        }
        %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
        %This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
        %%% let's omit above sentences completely. why ramble about monitors in the introduction? this ain't about model checking

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.\MP{ cit }

@@

The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness.\MP{cit}

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
%      MK: I don't think we need to mention this at all
This translation is considered {\em correct} if it is semantics--preserving.
%That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
Unfortunately, correct compilers may be insecure compilers\MK{citation needed}.
Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
For a compiler to be secure, this paper makes use of the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, consider \rul{properties $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} such that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}.
Then, \oul{$\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, the compiled version of $\src{\progvar}$, must robustly satisfy $\pi$}.

@@

This section presents the key meta-theoretic results of this paper and extends on prior work~\cite{kruse2022csc}.
A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
This is not intuitive in the sense that there are countless examples in the security domain where composition does not work.\MK{cite}
For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

Since the composition of secure compilers is again a secure compiler, the theorem generalizes to a whole chain of $n$ secure compilers.
Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8\MK{cit} eventually compiles in parts to $\irl{Ignition BC}$.
Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.

Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in the wild, there are countless instances of them.\MK{cit}
Consider the compilation passes of LLVM\MK{cit} whose source-code optimisations are industry--competitive.
In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
Take \gls{cf} and \gls{dce}.
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
However, this costs additional time until compilation is done, which can be a valuable resource\MK{cit}.
Hence, compiler engineers typically try to find an order of optimisations that yields well-optimised programs as fast as possible.
The following corollary justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

Besides the sequential composition, there are two other compositions that are more at the language- than the compiler--level.
Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
A practical example of this, which is, however, insecure, is {\tt Android Studio}\MK{cit}.
The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
The corresponding theorem is not listed, since there is no general definition of this kind of composition.

@@

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high--level overview.
The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.

>>>>

\begin{itemize}
  \item %
        %\Cref{sec:compprop} presents a formal framework to reason about compositions of security properties and demonstrate its usage with concrete examples:
        First, this paper studies the formalisation and the composition of security properties (\Cref{sec:compprop}) by focussing on those safety properties that are of interest for real-world compiler writers (as identified by the plethora of work enforcing such properties individually~\cite{akritidis2009baggy,nagarakatte2009soft,nagarakatte2010cets,dhumbumroong2020boundwarden,jung2021pico,nam2019framer,shankaranarayana2023tailcheck,younan2010paricheck,zhou2023fatptrs,bond2017vale,almeida2017jasmin,kuepper2023cryptopt,cauligi2019fact}\MK{even more cits?}).
        % 
        Starting from ways to formalise and enforce those properties individually, this paper shows how to compose both their formalisation and their composition.
        % 
        This paper then showcases the benefits of this composition by composing \gls{tms} and \gls{sms} into \gls{ms}, and ultimately adding \gls{cct} to the composition, yielding the the conjunction of \gls{ms} and \gls{cct}.
        % 
        The resulting security property is the golden standard of safety properties that secure cryptographic code should have~\MP{cits}.
        \MPin{
        	conclude that did this prop in a single pass, we show how to do it compositionally
        }
        %For each of these properties, we also give a monitor which runs on another trace model that excludes unneccessary information and show that their composition is behaviorally similar to the composition of the trace-based security properties.
        %This two-level approach enhances modularity: When working with the security properties, one only needs to reason about the corresponding monitor.
        %%% let's omit above sentences completely. why ramble about monitors in the introduction? this ain't about model checking

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.

@@

The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety or liveness~\cite{clarkson2008hyper}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
%For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.%\MP{not a fan of this sentence}
%This paper writes $\src{\progvar}$ and $\trg{\progvar}$ for programs written in $\src{L}$ and $\trg{L}$, respectively.
%      MK: I don't think we need to mention this at all
This translation is considered {\em correct} if it is semantics--preserving.
%That is, the compiler encodes any possible $\src{L}$--level feature in $\trg{L}$, such that the semantics are the same.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}.
Programs translated by insecure compilers can violate security properties that the programmer assumes to hold. %, given the presence of a context--level program.
For a compiler to be secure, this paper makes use of the robust compilation framework~\cite{abate2019jour}, which we summarize in the following definition.
For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, consider \rul{properties $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} such that \iul{the program $\src{\progvar}$ robustly satisfies $\pi$}.
Then, \oul{$\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, the compiled version of $\src{\progvar}$, must robustly satisfy $\pi$}.

@@

This section presents the key meta-theoretic results of this paper and extends on prior work~\cite{kruse2022csc}.
A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
This is not intuitive in the sense that there are a number of examples in the security domain where composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

Since the composition of secure compilers is again a secure compiler, the theorem generalizes to a whole chain of $n$ secure compilers.
Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.

Notably, real-world compilation chains also perform a series of passes whose main purpose is not necessarily to translate from one language to another, but to, e.g., optimise the code or enforce a certain property.
Both examples can be seen in the wild~\cite{nagarakatte2009soft,nagarakatte2010cets,akritidis2009baggy,wegman1991ccp,manjikian1997fusion}, it is infeasible to list them all.
Consider the compilation passes of LLVM\cite{lattner2004llvm} whose source-code optimisations are industry--competitive.
In LLVM, the pass manager can be used to schedule passes in a way where the optimisations yield the best results.
Take \gls{cf} and \gls{dce}.
Unfortunately, the order in which \gls{cf} and \gls{dce} are performed influences the final result of the compilation (see \Cref{fig:cfdceex}).
This {\em phase ordering problem} is well--known in literature and a practical solution is to simply perform a fixpoint iteration.
Compiler engineers typically try to find an order of optimisations that yields well--optimised programs for either code size~\cite{cooper1999geneticphases} or performance~\cite{kulkarni2006exhaustivephase}.
The following corollary justifies that any such order of compilation passes is valid with respect to security.
So, given two compilation passes \bul{$\cc[_{1}]{\trg{L}}{\trg{L}}$, $\cc[_{2}]{\trg{L}}{\trg{L}}$, both robustly preserving class $\class[_{1}]$ or $\class[_{2}]$, respectively}, \oul{for any order of their composition the composed compiler robustly preserves the intersection of $\class[_{1}]$ and $\class[_{2}]$}.

@@

Besides the sequential composition, there are two other compositions that are more at the language- than the compiler--level.
Nevertheless, consider $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
As an example, $\src{Java}$ and $\obj{Kotlin}$ compile to $\trg{JVM BC}$.
Technically speaking, they both lower to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language, or rather, intermediate representation, performs a kind of {\em upper} composition.
Given a secure compiler from $\src{Java}$ to $\trg{JVM BC}$ and one from $\obj{Kotlin}$ to $\trg{JVM BC}$, it is easy to envision that one can engineer a combined compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$.
A practical example of this, which is, however, insecure, is {\tt Android Studio}~\cite{androidstudio}.
The proof of \Cref{thm:rtp} can be easily adapted to also works for this {\em upper} composition.
The corresponding theorem is not listed, since there is no general definition of this kind of composition.

@@

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high--level overview.
The full abstraction approach~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.
